---
title: 'Blog Post Two'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(kableExtra)
library(lubridate)
```


```{r}
#loading data - I downloded from Rami Krispin's github and included the file in the repository. The data can also be found by installing the `coronavirus` package. 
load(file = "coronavirus.rda")
```


```{r ecuador}
#filtering data on Ecuador and pivoting the case types wider
ecuador <- coronavirus %>%
  filter(country=="Ecuador") %>%
  pivot_wider(names_from = type, values_from = cases)
```


## How has Ecuador's treatment of the coronavirus changed over time?

At the beginning of the year, the novel coronavirus (COVID-19) ravaged nations all around the word, in part, due to the world's inexperience in dealing with such a uniquely infectious disease. As time has gone on, however, nations have developed expertise in treating this disease. 

To gauge how Ecuador's ability to cope with the coronavirus has changed, let us look at how the number of deaths occuring within a week have changed. We will do this using data supplied by [Rami Krispin](https://github.com/RamiKrispin/coronavirus#readme)

```{r ecuador-death, fig.cap="On the x-axis is the week of the year and on the y-axis is the coronavirus-related deaths that occured during that week"}

ecuador_death_plot <- ecuador %>%
  mutate(week=week(date)) %>%
  group_by(week) %>%
  summarise(weekly_deaths=sum(death)) %>%
  ggplot() +
  geom_line(aes(x=week,y=weekly_deaths,color=weekly_deaths)) +
  scale_color_gradient(low = "green", high = "red") +
  xlab("Week of the Year") +
  ylab("Weekly Deaths") +
  theme(legend.title = element_blank()) +
  ggtitle("Coronavirus-related deaths in Ecuador")


ecuador_death_plot
```


As we can see, weekly deaths started off at zero before beginning a rapid rise through weeks 12-18 before peaking at week 19 with over 700 deaths. After that point, Ecuador seem to have gained experience in dealing with coronavirus as we then see a sharpish decline through week 26 after which we see the weekly deaths begin jumping up and down but generally following a steady decline.

Therefore, in answer to our question, we can say that Ecuador's treatment of coronavirus improved over time as the weekly deaths declined.


## Has Ecuador experienced more coronavirus-related death relative to its population than its neighbours?

Ecuador shares borders with Colombia and Peru. So, was Ecuador more or less affected by the coronavirus in terms of resulting death compared to its neighbours?
 
As we can see from the below table, Ecuador had considerably less total death due to the coronavirus relative to its neighbours. We can also see that it has a much lower population compared to its neighbours and as such its 'death rate' is much lower at 0.0126% compared to 0.0137% for Colombia and a whopping 0.0347% for Peru.

```{r deaths-table}
total_deaths <- coronavirus %>%
  filter(country==c("Ecuador","Peru","Colombia"),
         type=="death") %>%
  group_by(country) %>%
  summarise(total_death=sum(cases))

population <- tribble(
  ~country, ~population,
  "Colombia",50999546,
  "Ecuador",17701044,
  "Peru",33071098
)

countries_death_population <- inner_join(total_deaths,population,by="country") %>%
  mutate(death_rate_percentage=(total_death/population)*100) %>%
  arrange(-death_rate_percentage) %>%
  kable(caption="Ecuador, Peru and Colombia's coronavirus-related deaths relative to population") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

countries_death_population
```


## References

Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made
  Easy with lubridate. Journal of Statistical Software, 40(3),
  1-25. URL http://www.jstatsoft.org/v40/i03/.
  
Hao Zhu (2019). kableExtra: Construct Complex Table with 'kable'
  and Pipe Syntax. R package version 1.1.0.
  https://CRAN.R-project.org/package=kableExtra
  
Rami Krispin and Jarrett Byrnes (2020). coronavirus: The 2019
  Novel Coronavirus COVID-19 (2019-nCoV) Dataset. R package
  version 0.3.0. https://CRAN.R-project.org/package=coronavirus
  
Wickham et al., (2019). Welcome to the tidyverse. Journal of
  Open Source Software, 4(43), 1686,
  https://doi.org/10.21105/joss.01686
  
Worldometer.info., (2020). Countries in South America by population    (2020),
  https://www.worldometers.info/population
